### 2020-06-05 reviews/response.txt

We thank the reviewers for their exceptionally helpful and thorough
comments. Thorough reviews take time, interest, and expertise! Here,
we respond to reviewers' "Comments for Revision", marked by (C), and
due to limited space explain only a selection of detailed comments,
marked by (D). However, we have addressed all of the reviewers
comments in our revised paper either by making the requested change or
by providing additional explanation.

Our revision includes two substantive additions and numerous
improvements to explanations, graphics, and presentation. The first
substantive addition is a corroboration of our airflow implications on
temperature with temperature data from another publication. The second
substantive addition is a new result that job scheduling and folded
torus interconnect were contributing factors to temperature patterns,
and explain a peculiar interleaving column hazard variability.

We have added more explantion to most figures, making them more
self-contained and relevant to our conclusions. We also made some
small but useful tweaks in our narrative that is not highlited in blue
so as to not distract from the more substantive changes responding to
reviewer comments.

@@

(C)
We have strengthened our explanations in several sections and
particularly around our airflow explanation in Fig.4. Both substantive
improvements described above relate to this point, on one hand
strengthening it with temperature data, and on the other hand also finding
a more nuanced impact of scheduling.

(D)
Section 3: see changes in first paragraph.

Reviewer 3:

(C)
Generalization: This analysis is already generating interest from
Summit operations. A lesson learned is that Titan was not well
instrumented for data collection - thus our difficulty in obtaining
temperature data. Frontier will likely have even more
instrumentation. We now note this in our conclusions.

Novelty: We have added notes about novelty in the visualizations and
our analysis in the introduction.

### 2020-06-05 intro.tex
novel viz
novelty notes

We begin by describing related work in Sec.~\ref{section:related}. Our
data collection is described in Sec.~\ref{section:dataprep} and the
data cleaning, checking, and lifetime summarization process in
Sec.~\ref{section:dataclean}. A time between failures (TBF) analysis
is given in Sec.~\ref{section:tbf}. Survival analysis \rev{(SA)} based on
Kaplan-Meier modeling (KM) and Cox regression modeling (CR) of
relative hazard rates in various locations are in
Sec.~\ref{section:survival}. The goals of a TBF analysis and of
SA are different: TBF studies largely \rev{show} GPU
reliability affects the machine, whereas SA studies largely how the
machine affects GPU reliability.  Finally, we discuss our main
conclusions and recommendations in Sec.~\ref{section:discussion}.

>>>>

We begin by describing related work in Sec.~\ref{section:related}. Our
data collection is described in Sec.~\ref{section:dataprep} and the
data cleaning, checking, and lifetime summarization process in
Sec.~\ref{section:dataclean}. \rev{Here we develop novel
  visualizations of GPU life times and errors that were needed to
  understand the data and our data processing decisions.} A time
between failures (TBF) analysis is given in
Sec.~\ref{section:tbf}. Survival analysis \rev{(SA)} based on
Kaplan-Meier modeling (KM) and Cox regression modeling (CR) of
relative hazard rates in various locations are in
Sec.~\ref{section:survival}. The goals of a TBF analysis and of SA are
different: TBF studies largely \rev{show} GPU reliability affects the
machine, whereas SA studies largely how the machine affects GPU
reliability. \rev{The application of SA methods is novel in the HPC
  reliability context. It has roots in biostatistics and epidemiology,
  where it is used to discover causes of disease.} Finally, we discuss
our main conclusions and recommendations in
Sec.~\ref{section:discussion}.
